A Classical Density-Functional Theory for Describing Water Interfaces

### 2011-09-07 paper/project.tex
improve the dispersion section just a tad
Tested-on: bingley (Linux 2.6.32-5-amd64)

The dispersion free energy has the form~\cite{gil-villegas-1997-SAFT-VR}
\begin{align}
  F_\text{disp} &= A_1 + \beta A_2
\end{align}
where $A_1$ and $A_2$ are the first two terms in a high-temperature
perturbation expansion.  $A_1$ is the mean-field dispersion
interaction, which for a homogeneous fluid is given by
\begin{align}
  A_1 &= \frac12 n_b \int \varphi(\left|\xx\right|)
  g_{HS}(\left|\xx\right|) d\xx
\end{align}
The second dispersion term in the free energy $A_2$ describes the
effect of fluctuations resulting from the compression of the fluid due
to the dispersion interaction itself, and is commonly approximated
using the local compressibility approximation (LCA).

We use a free square-well function for the dispersion $\varphi$, which
has been demonstrated \textcolor{red}{(CITE)} to give a reasonable fit
to the equation of state.  Thus our model interaction has the form:
\begin{equation}
  \varphi(r) = \Theta(r-2 \lambda_\textit{disp} R)
\end{equation}
where $\Theta$ is the Heaviside step function.

\textcolor{red}{TODO:  Add the actual functions that we use and the
  actual form.  The above is really just an introduction.}

>>>>

The dispersion free energy has the form~\cite{gil-villegas-1997-SAFT-VR}
\begin{align}
  F_\text{disp} &= A_1 + \beta A_2
\end{align}
where $A_1$ and $A_2$ are the first two terms in a high-temperature
perturbation expansion.  $A_1$ is the mean-field dispersion
interaction, which reduces in the homogeneous limit to
\begin{align}
  \frac{A_1}{V} &= \frac12 n_b \int \varphi(\left|\xx\right|)
  g_{HS}(\left|\xx\right|) d\xx
\end{align}
The second dispersion term in the free energy $A_2$ describes the
effect of fluctuations resulting from the compression of the fluid due
to the dispersion interaction itself, and is commonly approximated
using the local compressibility approximation (LCA).

We use a free square-well function for the dispersion $\varphi$, which
is the choice used by Clark \emph{et al}~\cite{clark2006developing},
and allows a reasonable fit to the equation of state.  Thus our model
interaction has the form:
\begin{equation}
  \varphi(r) = \Theta(r-2 \lambda_\textit{disp} R)
\end{equation}
where $\Theta$ is the Heaviside step function.

The form of $A_1$ is given in
reference~\cite{gil-villegas-1997-SAFT-VR}, but is expressed in terms
of the filling fraction.  In order to apply this form to an
inhomogeneous density distribution, we construct an effective local
filling fraction $\eta{\textit{eff}}$, given by
\begin{align}
  \eta_{\textit{eff}}(\mathbf{r}) &= \frac{4\pi R^3}{3}
  \int \left(\frac{\pi}{\sigma}\right)^{\frac32}
  e^{-\frac{|\mathbf{r}-\mathbf{r}'|^2}{2\sigma^2}} n(\mathbf{r}') d\mathbf{r}'
\end{align}
This effective filling fraction is used throughout the dispersion
calculations.  This functional introduces an additional empirical
parameter $\sigma$ \fixme{(The above formula still needs fixing)}
which adjusts the length scale over which the dispersion interaction
is correlated.

\fixme{Do we want to include all the gory details? It's pretty
  complicated, which could be used to argue in either direction.  On
  the one hand, a complete summary that doesn't require users to look
  up the 1997 paper would be helpful.  On the other hand, these
  precise details are lengthy and a pain to write, and aren't all that
  interesting for most readers to read.}

The second term, $A_2$, which describes the contribution to the free
energy associated with fluctuations is given by
\begin{align}
  A_2 &= \int \frac12 \epsilon_\textit{disp}
              K_{HS}(\eta_{\textit{eff}}(\mathbf{r}))
              \eta_{\textit{eff}}(\mathbf{r})
              \frac{\delta A_1}{\delta \eta_\textit{eff}(\mathbf{r})}
                d\mathbf{r}
\end{align}
where $K_{HS}$ is the Percus-Yevick hard-sphere compressibility, given
by
\begin{align}
  K_{HS} &=
    \frac{(1 - \eta_\textit{eff})^4}{1 + 4(\eta_\textit{eff} + \eta_\textit{eff}^2)}
\end{align}
